The Dynamics of Anthropomorphic Effects in Human-Robot Interaction

### 2014-03-25 paper.tex
Moved back the section on Disruptive behavior to Sec Cognitive Correlates
This discussion on disrptive behavior actually relies on the cognitive
interpretations of anthropomorphism, so better have it there.
Tuning of the model of the Dynamics + {expectation mismatch -> attractiveness delta}

% for use inside a math environment
\newcommand{\ant}[1][]{%
      \ifthenelse{\isempty{#1}}%
        {\mathcal{A}}
        {\mathcal{A}(#1)}
}
\newcommand{\ICA}{{$\mathcal{A}_0$~}}
\newcommand{\SLA}{{$\mathcal{A}_\infty$~}}
\newcommand{\sla}{{\mathcal{A}_\infty}}
\newcommand{\AntMax}{{$\mathcal{A}_{max}$~}}

@@

The phenomenological model of anthropomorphism we propose
(Figure~\ref{fig:dynamics},~\cite{lemaignan2014dynamics}) represents how the
level of anthropomorphic effects (\ie observable manifestations of
anthropomorphism, as defined in section~\ref{sec:intro}) evolves over a
long-term human-robot interaction. By long-term interaction, we mean direct
(non-mediated), repeated interaction with the same robot, over a long period of
time.

While the model primarily considers \emph{long-term interaction} -- direct
(non-mediated), repeated interaction with the same robot, over an extended
period of time (typically longer than a week) --, we also formally introduce a
so-called \emph{novelty effect}~\cite{kanda_interactive_2004} that models the
first phase of human-robot interaction, during which a specific increase of
anthropomorphic interactions is observed.

In this model, anthropomorphism is quantified by a \emph{normalized level of
anthropomorphic effects}: because anthropomorphic effects are not quantified on
an absolute scale, we present them as a normalized value, that spans from a
minimum (no anthropomorphic effects: $\ant=0$) to a maximum (noted \AntMax, with
$t_{max} = \operatorname{arg\,max}_t \, \ant[t]$). We also discuss at
section~\ref{measuring} possible methodologies for the quantitative and
qualitative assessment of anthropomorphic effects.

@@

% vertical axis
\draw[->] (0,0) -- (0,6) node[anchor=east] {};
\draw(-0.8,3) node[rotate=90,anchor=south] {\scriptsize Normalized level of anthropomorphic effects};

@@

\draw[dotted] (0, 5) -- (6.2,5);
\draw[dotted] (0, 3) -- (6.2,3);
\draw[<->] (6.1,3) -- (6.1,5) node[anchor=east] {};
\draw (6.2,4) node[rotate=90, anchor=north, align=center] {\tiny Expectation\\\tiny mismatch};

@@

@@

@@

\paragraph{Expectation mismatch}

Besides, we propose to call \emph{expectation mismatch} the delta between the
initial capital of anthropomorphism for a given tuple (user, robot, context)
and the anthropomorphism peak of novelty effect (Figure~\ref{fig:dynamics}).
\fixme{Need to rediscuss what this delta actually represent...}

@@

\subsection{Relation to the model of anthropomorphism}

These cognitive levels overlap but do not exactly match the
\emph{Initialization}, \emph{Familiarization} and \emph{Stabilization} phases
introduced in our model of the dynamics of anthropomorphism. In particular,
the cognitive levels 1 and 2 are both included in the \emph{initialization} phase
of the anthropomorphism model. Sub-cognitive anthropomorphism typically
\emph{initiates} the novelty effect by rapidly engaging the human in the
interaction through an initial projected agency, whereas cognitive level 2
(projection of familiar mental models) supports the novelty effect by inducing
beliefs that the robot is set up with possibly complex cognitive abilities.

@@

>>>>

\newcommand{\AntNorm}[1][]{%
      \ifthenelse{\isempty{#1}}%
      {$\widehat{\mathcal{A}}$}
      {$\widehat{\mathcal{A}}(#1)$}
}

% for use inside a math environment
\newcommand{\ant}[1][]{%
      \ifthenelse{\isempty{#1}}%
        {\mathcal{A}}
        {\mathcal{A}(#1)}
}

\newcommand{\antNorm}[1][]{%
      \ifthenelse{\isempty{#1}}%
      {\widehat{\mathcal{A}}}
      {\widehat{\mathcal{A}}(#1)}
}

\newcommand{\ICA}{{$\mathcal{A}_0$~}}
\newcommand{\SLA}{{$\mathcal{A}_\infty$~}}
\newcommand{\sla}{{\mathcal{A}_\infty}}
\newcommand{\AntMax}{{$\mathcal{A}_{max}$~}}
\newcommand{\antMax}{{\mathcal{A}_{max}}}

@@

The phenomenological model of anthropomorphism we propose
(Figure~\ref{fig:dynamics}~\citep{lemaignan2014dynamics}) represents how the
level of anthropomorphic effects~\Ant[R,H,C,t] (\ie observable manifestations of
anthropomorphism, as defined in section~\ref{sec:intro}) evolves over a
long-term human-robot interaction.

While the model primarily considers \emph{long-term interaction} -- direct
(non-mediated), repeated interaction with the same robot, over an extended
period of time (typically longer than a week), we also formally introduce a
so-called \emph{novelty effect}~\cite{kanda_interactive_2004} that models the
first phase of human-robot interaction, during which a specific increase of
anthropomorphic interactions is observed.

In this model, anthropomorphism is quantified by a \emph{normalized level of
anthropomorphic effects} $\antNorm[R,H,C,t] = \frac{\ant[R,H,C,t]}{\antMax}$: because anthropomorphic effects are
not quantified on an absolute scale, we present them as a normalized value, that
spans from a minimum (no anthropomorphic effects: $\ant=0$) to a maximum (noted
\AntMax, with $t_{max} = \operatorname{arg\,max}_t \, \ant[t]$). For readability
reasons, we however omit the hat on \Ant in the following sections. We discuss
at section~\ref{measuring} possible methodologies for the quantitative and
qualitative assessment of anthropomorphic effects.

@@

% vertical axis
\draw[->] (0,0) -- (0,6) node[anchor=east] {};
\draw(-0.8,3) node[rotate=90,anchor=south] {\scriptsize Anthropomorphic effects};

@@

\draw[dashed] (0.6, 0) -- (0.6,5.1);
\draw (0.6,0) node[anchor=north] {$t_{max}$};

\draw[dotted] (0, 5) -- (6.2,5);
\draw[dotted] (0, 3) -- (6.2,3);
\draw[<->] (6.1,3) -- (6.1,5) node[sloped, above, midway] {$\Delta_{a}$};

@@

\fxfatal{Julia, do you have smthg to provide an initial account of what the
novelty effect is?}

@@

\paragraph{Attractiveness Delta}

Besides, we propose to call \emph{attractiveness delta}, noted $\Delta_{a}$, the
difference between the initial capital of anthropomorphism \ICA for a given tuple (user,
robot, context) and the anthropomorphism peak \AntMax (Figure~\ref{fig:dynamics}).
This can be interpreted as the initial \emph{excitement} over the robot, or how
much a user tend to increase his/her anthropomorphic engagement when he/she
starts to actually interact with the robot.

@@

@@

\subsection{Relation to the dynamics of anthropomorphism}

These cognitive levels overlap over time but do not exactly match the
\emph{Initialization}, \emph{Familiarization} and \emph{Stabilization} phases
introduced in our model of the dynamics of anthropomorphism. In particular,
the cognitive levels 1 and 2 are both included in the \emph{initialization} phase
of the anthropomorphism model. Sub-cognitive anthropomorphism typically
\emph{initiates} the novelty effect by rapidly engaging the human in the
interaction through an initial projected agency, whereas cognitive level 2
(projection of familiar mental models) supports the novelty effect by inducing
beliefs that the robot is set up with possibly complex cognitive abilities.

@@

A cognitive interpretation of anthropomorphism also allows to better interpret
the role of unexpected robot behaviors, that we are \emph{disruptive} in regard
to the cognitive process of building a mental model of the robot.
